### 2017-10-11 OaklandSubmission/related.tex
1.1

\section{Related work}
\label{sec:related}

\divya{Related Work: writing some citations here to be used appropriately.}
\begin{enumerate}
\item Secure Computation of MIPS Machine Code https://eprint.iacr.org/2015/547
\item TASTY https://eprint.iacr.org/2010/365.pdf
\item ABY cites many papers on first page of intro for using a mix of HE and boolean MPC; but they didnt give much gain because HE is expensive

\end{enumerate}