### 2017-10-13 OaklandSubmission/related.tex
related

\section{Related work}
\label{sec:related}

\divya{Related Work: writing some citations here to be used appropriately.}
\begin{enumerate}
\item Secure Computation of MIPS Machine Code https://eprint.iacr.org/2015/547
\item TASTY https://eprint.iacr.org/2010/365.pdf. Does not have language; user has to know what parts in HE and what in Yao and writes something similar to ABY
\item ABY cites many papers on first page of intro for using a mix of HE and boolean MPC; but they didnt give much gain because HE is expensive
\nc{I will look into this and write.}
\item verilog, hardware tool chain paper http://encrypto.de/papers/DDKSSZ15.pdf
\item automatic choosing between HE and Yao https://eprint.iacr.org/2014/200.pdf 
I think this one still does not have a language; I don't know how to specify function

\end{enumerate}