### 2017-10-26 OaklandSubmission/impl.tex
impl

\section{Implementation}
\label{sec:impl}
We discuss some  implementation details of \tool.
The  \tool compiler is written in Python and
compiles each of our benchmarks in under a second to C++ code that makes calls to the ABY
library~\cite{aby}. We set the security paramenter to 128 and use OT extension-based arithmetic multiplication triples generation. ABY provides facilities to use multiple threads\footnote{The protocol has an offline phase and an online phase. Although multiple threads can be used in the offline phase, the online phase is single threaded.} and we use at most four threads in our evaluation.

The current language implementation supports the following operators:
addition, subtraction, multiplication, division by powers of two, left shift,
logical and arithmetic right shifts, bitwise-(and, or, xor), unary negation,
bitwise-negation, logical-(not, and, or, xor), and comparisons (less than, greater than, equality).
Because of their high cost, integral division and floating-point operators are not supported natively by \tool.
However, we have implemented integral division in 30 lines of \tool and floating-point support in
ABY is under active development~\cite{ddkssz15}.

Some of our benchamrks require accessing arrays at secret locations.
For example, consider the expression $A[x]$ where $A$ is an array of size 2 and $x$ is shared.
\tool enforces the array indices to be public as accessing arrays at secret locations is expensive.
However, the developer can still express this functionality in \tool
as $x > 0 ? A[1] : A[0]$. In general, a secret access to an array of size $2^n$ requires $2^{n}-1$ multiplexers in \tool.

We provide implementation details of compilation rules {\sc T-Read} and {\sc T-Decl} (Figure~\ref{fig:compile}).
In particular, the check $\Gamma\vDash e < n$ in {\sc T-Read}
which prove that the array indices are appropriately bounded
is discharged by an off-the-shelf solver (SeaHorn~\cite{seahorn}).
The solver takes less than a minute on our largest benchmark.
Our implementation of {\sc T-Decl} assigns type labels conservatively.
Only the variables that govern the control flow, i.e., variables
in the predicates of if-conditions and loop counters of for-loops are assigned public labels.
All other variables are  assigned arithmetic labels (and can later be coerced to boolean).
In the future, we would like to explore more intelligent strategies for label assignment.

Finally, the compilation rules in Figure~\ref{fig:compile} can introduce redundant coercions from arithmetic to
boolean  and vice versa.
The  compiler performs several optimizations to reduce these.
For example, on each coercion, \tool stores the pair of arithmetic share and boolean share involved in the coercion.
\tool invalidates such pairs when the variables corresponding to the shares are overwritten by assignments.
In subsequent coercions, \tool  reuses valid pairs (if available) instead of recomputing them afresh.