### 2017-10-27 OaklandSubmission/pipe.tex
partitioning

\section{Secure code partitioning}
\label{sec:pipe}

We  describe ``secure code partitioning'' that allows us to execute {\it large} programs, i.e., programs with a large number of operations.
In more detail, let $P$ be a program in our source language that generates a circuit $\crct$. For a large enough $P$, the circuit $\crct$  can be larger than the memory size\footnote{In fact, there is an upper limit of $2^{32}-1$ gates for the circuit size in ABY but for most machines the memory limit is hit first} and fail to execute. Partitioning enables us to
execute such programs via a source to source transformation that is oblivious to the underlying \mpc backend. Partitioning decomposes the program $P$ into a sequence of smaller \tool programs $Q_1||Q_2||\cdots||Q_k$ (defined below). We compile and execute each $Q_i$ sequentially, feeding the outputs of $Q_i$ as state information to $Q_{i+1}$. We prove that partioning is correct ($P$ and $Q_1||Q_2||\cdots||Q_k$ compute the same functionality) and secure (execution of $Q_1||Q_2||\cdots||Q_k$ does not reveal any more information than $P$). More details follow.

Let $P$ be a program that takes (secret) inputs $x$ and $y$ and produces an output $z$ to both parties (Alice and Bob). Let $P_1||P_2||\cdots||P_k$ be a sequence of programs such that the following holds: Define $s_0 = \bot$ (the public empty state). For all $1\leq i\leq k-1$, $P_i$ takes inputs $x, y$ and $s_{i-1}$ and outputs  state information $s_i$. Finally, $P_k$ takes inputs $x,y$ and $s_{k-1}$ to  output $z$. Each program $P_i$ ($1\leq i\leq k$) is required to be ``sufficiently'' small (as described later). It is possible to decompose any program $P$  into such $P_1||P_2||\cdots||P_k$. If \tool generates circuit $\crct_i$ from $P_i$, the parties can execute $\crct_1||\crct_2||\cdots||\crct_k$ sequentially (in a distributed setting)  to obtain $s_1,\cdots,s_{k-1},$ and finally output $z$. At the $i^{\textrm{\tiny{th}}}$ step, the parties only need to store information proportional to $x,y,s_{i-1}$ and $\crct_i$ (which is much smaller than $\crct$). However, this execution enables the parties to learn $s_i$ (for all $1\leq i\leq k-1$), which completely breaks the security.

To overcome this problem, we define a sequence of new programs $Q_i$ ($1\leq i\leq k$) as follows. Once again, define $s_0 = \bot$. Without loss of generality, let all $s_i$ be values in some additive ring $(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$ i.e., the additive ring of integers modulo $2^{64}$). 
Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all $r_i$ values are generated by a pseudorandom function). Let $Q_1$ be the program that takes as inputs $x,y, r_1$ (and empty state $s_0$), runs $P_1$ (as defined above) to get $s_1$ and then outputs $o_1 = s_1 + r_1$ {\em only to} Bob\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output. To do so, we let $Q_1$ output $o_1 = s_1+r_1+t_1$ (where $t_1$ is random and chosen by Bob) and have both parties learn this. We can then appropriately modify the remaining steps as well.}. Alice's output from $Q_1$ is $r_1$. Next, every $Q_i$ ($2\leq i\leq k-1$) takes as inputs $x,y,r_{i-1},o_{i-1},r_i$, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$ only to Bob. Again, Alice's output from $P_i$ is $r_i$. The last program $Q_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $P_k$ on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both parties.

Given $P$, partitioning first decomposes $P$ into $P_1||P_2||\cdots||P_k$ with the guarantee that the circuits generated for every $Q_i$ defined above are small enough to fit in the machine memory. We then compile and execute the $k$ programs $Q_1||Q_2||\cdots||Q_k$ sequentially (freeing up memory usage after  execution of each $Q_i$) to finally output $z$ to both parties.

\begin{theorem}[Correctness and security of partitioning]
Let $P$ be a program in our source language with output $z$, that is decomposed into programs $Q_1||Q_2||\cdots ||Q_k$ (as defined in the partioning procedure above) and compiled into the protocol $\prot_1||\prot_2\cdots ||\prot_k$ that outputs $r_1||r_2||\cdots||z$ (to Alice) and $o_1||o_2||\cdots||z$ (to Bob), then $\prot$ securely realizes $P$.  
\end{theorem}

\noindent {\em Proof.} From Theorem \ref{theorem:security}, it follows that $\prot_i$ securely realizes $Q_i$ for every $1\leq i\leq k$. Hence, we know that the only information learnt by Alice is $r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is $o_1||o_2||\cdots||o_{k-1}||z$. Since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, the proof follows.

Our current implementation of secure code partitioning requires a manual step.
In particular, the developer needs to decompose the large program $P$ into 
the sequence of small programs $P_1||\ldots||P_k$ manually.
Then \tool generates $Q_1||\ldots||Q_k$ automatically.
Automating this decomposition step requires an analysis that can 
statically estimate the resource usage of a \tool program and we leave the construction of such
an analysis as future work.