### 2017-10-27 OaklandSubmission/security.tex

\subsubsection*{Security theorem} The protocols we generate provide
simulation-based security against a semi-honest adversary, in the
framework of ~\cite{gmw,can00,can01}. At a very high level, in this
framework, parties are modeled as non-uniform interactive turing
machines (ITMs), with inputs provided by an environment $\env$. An
adversary $\adv$, selects and ``corrupts'' one of the parties -
however, $\adv$ still follows the protocol specification. $\adv$
interacts with $\env$ that observes the view of the corrupted
party. At the end of the interaction, $\env$ outputs a single bit based on the output of the honest party and the view of the adversary. Two
different interactions are defined: the {\em real world} and an {\em
  ideal world}. In the real interaction, the parties run the protocol
$\prot$ in the presence of $\adv$ and $\env$. Let
$\real_{\prot,\adv,\env}$ denote the  distribution ensemble
describing $\env$'s output in this interaction. 
In the ideal interaction, parties send their inputs to  a
trusted functionality that performs the desired
computation truthfully. Let $\simu$ (the simulator) denote the
adversary in this ideal execution, and $\ideal_{\F,\simu,\env}$
the distribution ensemble describing $\env$'s output after
interacting with the ideal adversary $\simu$. A
protocol $\prot$ is said to {\em securely realize} a functionality
$\F$ if for every adversary $\adv$ in the real interaction, there is
an adversary $\simu$ in the ideal interaction, such that no
environment $\env$, on any input, can tell the real interaction apart
from the ideal interaction, except with negligible probability (in the
security parameter $\secparam$). More precisely, the above two
distribution ensembles  are computationally indistinguishable.

We shall assume a cryptographic \mpc backend that securely implements any circuit $\crct$ that is output by our compiler (see \figureref{circuits}). In more detail, this means that for any well-typed source program $s$, let $\crct$ be the circuit output by our compiler (as in \theoremref{correctness}). We assume that there exists a \mpc protocol $\prot$ that securely realizes the functionality $\crct$ and let $\simu_{\tiny{2pc}}$ be the corresponding simulator  (that runs on $\crct$, input of the corrupt party and the output obtained from trusted functionality for $\crct$). We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\crct$ output by our compiler\footnote{\divya{We note that \cite{aby} provides a \mpc protocol for a more general .....}}. We are now ready to state and prove our security theorem.

\begin{theorem}[Security]\label{theorem:security}
Let $s$ be any functionality or program in our source language with outputs (or observations) $O_1$, that is compiled into a circuit $\chi$ (as defined in Theorem \ref{theorem:correctness}). Let protocol $\prot$ be the two-party secure computation protocol that securely realizes $\chi$ (as defined above). Then, $\prot$ securely realizes $s$.  
\end{theorem}

\noindent {\em Proof.} 
Our simulator $\simu$ simply runs our compiler on program $s$ to obtain $\crct$. It is crucial that this compilation to circuits does not require the secret inputs of the parties. Next, $\simu$ sends the input of the corrupt party to the trusted functionality of $s$ to obtain outputs $O_1$. Note that $O_1$ is same as the observations in the source semantics. 
By \theoremref{correctness}, these outputs $O_1$ are identical to outputs (or observations) $O_2$ of $\crct$ under circuit semantics. 
Next, $\simu$ runs $\simu_{\tiny{2pc}}$ on $\crct$, input of the corrupt party and $O_2$. 
From the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ is indistinguishable from the real view of the $\prot$ for $\crct$ between the two parties. Hence, the security follows.

% and upon receiving output $O_1$, executes $\simu_{\tiny{2pc}}$ on $\chi$ and $O_1$. Let $O_2$ be the outputs (or observations) of $\chi$. First, from Theorem \ref{theorem:correctness}, we have that the observations in $s$ and $\chi$ are the same (i.e. $O_2 = O_1$). Now, from the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ (with circuit $\chi$ and output $O_2$) is indistinguishable from the real view of the $\prot$ (when executed on circuit $\chi$ and when the output is $O_2$). Combining these two statements, the proof of the theorem follows. 
\aseem{
  Also, we should emaphasize here that the ideal semantics is the
  source semantics from Figure x and the real semantics is the circuit
  semantics from Figure y. For the target language semantics, we
  should emphasize that it does not have access to the secrets, so
  there are no observations there.
}

\nc{OK now?}
  %% \\
  %% \[
  %% \begin{array}{rrcl}
  %%   \ftext{Value} & v &::=& c \mid [\overline{c_{i}}]_{n}\\
  %%   \ftext{Observation} & O &::=& \cdot \mid c; O\\
  %% \end{array}
  %% \]