### 2017-10-30 OaklandSubmission/conclude.tex
cutting

\section{Conclusion}
\label{sec:conclude}
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols from high-level programs.
The generated protocols comprise combinations
of arithmetic and boolean circuits and have performance comparable to
or better than the custom, specialized
protocols from previous works