### 2017-10-30 OaklandSubmission/pipe.tex
pipe

\section{Secure code partitioning}
\label{sec:pipe}

\newcommand{\prog}{s}
\newcommand{\progn}{t}
\newcommand{\seq}{||}
\newcommand{\state}{q}

We  now describe our ``secure code partitioning'' technique that
allows \tool to execute programs that require large circuits.
Let $\prog$ be a program in our source language that
generates a circuit $\crct$. For some programs, the circuit $\crct$
can be larger than
the memory size\footnote{In fact, there is
  an upper limit of $2^{32}-1$ gates for the circuit size in ABY but
  for most machines the memory limit is hit first.} and fail to
execute. Partitioning enables us to 
execute such programs via a source to source transformation that is
oblivious to the underlying \mpc backend. Partitioning decomposes the
program $\prog$ into a sequence of smaller \tool programs
$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ (as defined below), such
that the circuit size
requirement for each of the $\prog_i$ itself is manageable. We compile
and execute each $\prog_i$ sequentially, feeding the outputs of
$\prog_i$ as state
information to $\prog_{i+1}$. We prove our partitioning scheme to be
correct ($\prog$ and $\prog_1\seq\prog_2\seq\dots\seq\prog_k$ compute
the same functionality) and secure (execution of
$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ does
not reveal any more information than $s$). More details follow.

Let $\prog$ be a program that takes (secret) inputs $x$ from Alice and
$y$ from Bob and produces an output $z$ to both parties. Let
$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ be a sequence of programs
such that the following holds. Define $\state_0 = \bot$ (the public empty
state). For all
$1\leq i\leq k-1$, $P_i$ takes inputs $x, y$ and $\state_{i-1}$ and
outputs state $\state_i$. Finally, $\prog_k$ takes inputs $x, y$ and
$\state_{k-1}$ to  output $z$. It is
possible to decompose any program $\prog$  into such
$\prog_1\seq\prog_2\seq\dots\seq\prog_k$. If \tool generates circuit
$\crct_i$ from
$\prog_i$, the parties can execute
$\crct_1, \crct_2, \dots \crct_k$
sequentially (in a distributed setting)  to obtain
$\state_1,\dots,\state_{k-1},$ and finally output $z$. At the
$i^{\textrm{\tiny{th}}}$ step, the parties only need to store
information proportional to $x,y,\state_{i-1}$ and $\crct_i$ (which is
much smaller than $\crct$). However, this execution enables the
parties to learn $\state_i$ (for all $1\leq i\leq k-1$), which
completely breaks the security.

To overcome this problem, we define a sequence of new programs
$\prog'_i$ ($1\leq i\leq k$) as follows. Once again, define $\state_0
= \bot$. Without
loss of generality, let all $\state_i$ be values in some additive ring
$(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$,
i.e., the additive ring of integers modulo $2^{64}$).
Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from
the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all
$r_i$ values are generated by a pseudorandom function). Let $\progn_1$
be the program that takes as input $x,r_1$ from Alice and $y$ from Bob
(and empty state $\state_0$), and runs $\prog_1$ (as defined above) to
compute $\state_1$ and then outputs $o_1 = \state_1 + r_1$ {\em only
  to} Bob\footnote{While the description of our protocol here assumes
  that
  the underlying backend supports only one party receiving output,
  this is only a simplifying assumption, and we can easily modify our
  protocol in the case where both parties must receive the same
  output. To do so, we modify $\progn_1$ to output $o_1 = s_1+r_1+r'_1$ to
  both parties (where $r'_1$ is random and chosen by Bob). We can then
  appropriately modify the remaining steps as well.}. Alice's output
from $\progn_1$ is $r_1$. Next, every $\progn_i$ ($2\leq i\leq k-1$)
takes as inputs $x,r_{i-1},r_i$ from Alice and $y,o_{i-1}$ from Bob,
runs $\prog_i$
on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes
subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$
only to Bob and $r_i$ only to Alice. The last
program $\progn_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both
parties.

Given $\prog$, partitioning first decomposes $\prog$ into
$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ with the guarantee that the
circuits generated for every $\prog'_i$, as defined above, are small
enough to fit in the machine memory. We then compile and execute the
$k$ programs $\progn_1,\progn_2,\ldots,\progn_k$ sequentially
(freeing up memory usage after execution of each $\progn_i$) to
finally output $z$ to both parties.

\begin{theorem}[Correctness and security of partitioning]
If $\prog_1||\prog_2||\ldots||\prog_k$ is a decomposition of a program $\prog$, then there exists a sequence of programs $\progn_1, \progn_2, \ldots, \progn_k$ and protocols $\prot_1, \prot_2, \ldots, \prot_k$ such that $\prot_i$ securely realizes $\progn_i$ and  $\prot
= \prot_1\seq\prot_2\dots\seq\prot_k$ securely realizes $s$.
\end{theorem}

\noindent {\em Proof.} Let $\progn_1, \ldots, \progn_k$ be the sequence of programs as defined above corresponding to the decomposition $\prog = \prog_1||\prog_2||\ldots||\prog_k$. \divya{Do we say that $\progn_i$ is a well-typed?} For every $1\leq i\leq k$, let $\prot_i$ be the \mpc protocol output by our framework for $\progn_i$. Then, by \theoremref{correctness},\theoremref{security}, $\prot_i$ securely realizes $\progn_i$. That is, for every $1\leq i \leq k-1$, the protocol $\prot_i$ provides observations $..$ to Alice and $...$ to Bob.

\begin{theorem}[Correctness and security of partitioning]
Let $\prog$ be a program in our source language with output $z$, that
is decomposed into programs
$\prog'_1\seq\prog'_2\seq\dots\seq\prog'_k$ (as defined in
the partioning procedure above) and compiled into the protocol $\prot
= \prot_1\seq\prot_2\dots\seq\prot_k$ that outputs $r_1\seq
r_2\seq\dots\seq z$
(to Alice) and $o_1\seq o_2\seq\dots||z$ (to Bob), then $\prot$ securely
realizes $P$.
\end{theorem}

\noindent {\em Proof.} From \theoremref{security}, it follows that
$\prot_i$ securely realizes $\prog'_i$ for every $1\leq i\leq
k$. Hence, we know that the only information learnt by Alice is
$r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is
$o_1||o_2||\cdots||o_{k-1}||z$. Since $r_i$ and $o_i$ ($1\leq i\leq
k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs
received by the adversary can be easily simulated given the final
output $z$.

Our current implementation of secure code partitioning requires a manual step.
In particular, the developer needs to decompose the large program $\prog$ into 
the sequence of small programs $\prog_1\seq\dots\seq\prog_k$ manually.
Then \tool generates $\prog'_1\seq\dots\seq\prog'_k$ automatically.
Automating this decomposition step requires an analysis that can 
statically estimate the resource usage of a \tool program and we leave
the construction of such an analysis as future work.