### 2017-10-30 OaklandSubmission/main.tex

\documentclass[conference,compsoc]{IEEEtran}

\input{headers}

\begin{document}
\title{\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
% \author{\IEEEauthorblockN{Michael Shell}
% \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
% Georgia Institute of Technology\\
% Atlanta, Georgia 30332--0250\\
% Email: http://www.michaelshell.org/contact.html}
% \and
% \IEEEauthorblockN{Homer Simpson}
% \IEEEauthorblockA{Twentieth Century Fox\\
% Springfield, USA\\
% Email: homer@thesimpsons.com}
% \and
% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
% \IEEEauthorblockA{Starfleet Academy\\
% San Francisco, California 96678-2391\\
% Telephone: (800) 555--1212\\
% Fax: (888) 555--1212}}

\maketitle
\begin{abstract}
Secure two-party computation (\mpc) allows two mutually distrusting parties to compute any function on their joint inputs, without revealing anything other than the output of the function to one another. %While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
% Prior work in designing such protocols, suffered from either a) a lack of programmability (e.g. requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning classification tasks); or c) required cryptographers to design hand-crafted protocols for various applications.
 In this work, we present \tool: a 
%new ``cryptographic-cost aware'' 
\mpc framework that generates efficient and scalable \mpc protocols from high-level programs that 
%contain absolutely no cryptographic details. Programs in our language 
are easy-to-write and are statically verified. 
%Furthermore, 
The \mpc protocols generated by our framework match or outperform recent works that provide hand-crafted \mpc protocols for various functionalities such as classification and matrix factorization (up to a 19x factor!).
\end{abstract}
\IEEEpeerreviewmaketitle

\graphicspath{{./Images/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\input{intro}
%\input{fig-ex-sml}
%\input{fig-ex-aby}

\input{ex}
\input{ld}
\input{security}
\input{pipe}
\input{impl}

\input{eval}
\input{train}
\input{related}
\input{conclude}

%\newpage
\bibliographystyle{abbrv}
\bibliography{bib}
%\balance

\appendix
\input{app-stash}

\end{document}