### 2017-10-31 OaklandSubmission/main.tex
authors

\documentclass[conference,compsoc]{IEEEtran}

\input{headers}

\begin{document}
\title{\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
% \author{\IEEEauthorblockN{Nishanth Chandran \hspace{10pt} Divya Gupta  \hspace{10pt} Aseem Rastogi \hspace{10pt}  Rahul Sharma \hspace{10pt} Shardul Tripathi }
% \IEEEauthorblockA{Microsoft Research India}
%}

\maketitle
\begin{abstract}
  %% Secure two-party computation (\mpc) allows two mutually distrusting
  %% parties to compute any function on their joint inputs, without
  %% revealing anything other than the output of the function to one
  %% another.

  %While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
% Prior work in designing such protocols, suffered from either a) a lack of programmability (e.g. requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning classification tasks); or c) required cryptographers to design hand-crafted protocols for various applications.
We present \tool: a 
%new ``cryptographic-cost aware'' 
secure two-party computation (\mpc) framework that generates efficient
\mpc protocols from high-level, easy-to-write, programs.
\tool provides formal correctness and security guarantees while maintaining performance and scalability.
%% that 
%% %contain absolutely no cryptographic details. Programs in our language 
%% are easy-to-write and are statically verified. 
%Furthermore,
Previous language frameworks such as CBMC-GC, ObliVM, SMCL, Wysteria, etc. generate protocols that use either arithmetic or boolean circuits exclusively. Our compiler is the first to generate protocols that combine both arithmetic sharing and garbled circuits  for better performance.
%We give the first compiler that generates protocols which combine both arithmetic and boolean circuits
%for better performance compared to protocols that use either arithmetic or boolean circuits exclusively.
We
empirically demonstrate that the
protocols generated by our framework match or outperform (up to
19x)
recent works that provide hand-crafted protocols for various
functionalities such as secure  prediction  and matrix factorization.
%\tool provides formal correctness and security guarantees.
%% (up to
%% a 19x factor!).
\end{abstract}
\IEEEpeerreviewmaketitle

\graphicspath{{./Images/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\input{intro}
%\input{fig-ex-sml}
%\input{fig-ex-aby}

\input{ex}
\input{ld}
\input{security}
\input{pipe}
\input{impl}

\input{eval}
\input{train}
\input{related}
\input{conclude}

%\newpage
\bibliographystyle{abbrv}
\bibliography{bib}
%\balance

%\appendix
%\input{app-stash}

\end{document}