### 2018-02-07 usenix18/appendix_b.tex
appendix a

\newpage

\section{Example of secure code partitioning}