### 2018-03-23 usenix18/rebuttal.txt
work

We thank the reviewers for their comments. We structure the rebuttal
to address high-level concerns in the first ~500 words, addressing
remaining concerns in a detailed section.

Comparison with Yao-only approach, Groce et al., and CBMC-GC etc.:

We are not the first to observe that the Yao-only approach does not
scale well beyond small examples ([18], [26], [31], [40],
[37]), especially for large-scale arithmetic computations commonly
found in Machine Learning (ML) algorithms -- the main use case of
EzPC. Indeed, this observation is one of the main driver behind
development of specialized MPC protocols in previous work ([40],
[37]). Therefore, it is critical to combine the  arithmetic and
boolean circuits for the kind of ML algorithms we apply EzPC to.

We thank Reviewer A for the Groce et al. reference. We present a
thorough comparison in the detailed section below. At a high-level,
the two overlapping benchmarks in Groce et al. are quite small, and
so, it is not surprising that the Yao-only approach works well. It is
not clear that Groce et al.'s approach would scale to large EzPC
examples (CIFAR 10 or matrix factorization). Furthermore, Groce et
al. improve the "online" protocol time at the cost of more "offline"
computation, while we are interested in minimizing the overall time
(more below).

As for comparing with other compilers, we compare EzPC with the best
known previous work, except for the benchmarks such as Bonsai, where
EzPC is the first MPC implementation. Compilers such as
CBMC-GC have not been used at the scale of benchmarks that we have
presented in the paper.

Performance gains:

For large benchmarks, such as CIFAR 10 (Table 7) and Matrix
Factorization (Table 8), EzPC performs much better than the previous
best known algorithms (~2x and ~19x resp.). Unsurprisingly, for small
benchmarks in Tables 1-4, where the baseline time itself is only
~5seconds, the performance gains of EzPC do not seem as impressive.

Language features:

We agree with the reviewers that the EzPC language itself is quite
simple, and currently does not support features such as secret
indices, and public inputs. However, the EzPC design has been driven
by machine learning, specifically secure prediction case studies. In
our experience, we have found that even with such a simple language,
we can implement a lot of such examples.

We also remark that while the EzPC language is simple, its compiler is
the first compiler that generates a mix of arithmetic and boolean
circuits, while hiding all the low-level crypto details from the
programmer.

Cost aware compiler:

As mentioned in Sec. 5, our compiler currently applies certain
heuristics to choose arithmetic or boolean circuit for different parts
of the program. Even with such simple heuristics, we are able to
achieve the demonstrated efficiency and scalability. A more precise
cost model that factors in the cost of network, input sizes, etc. is
the immediate future work for us.

==============================================

Detailed responses:

Comparison with Groce et al.:

Groce et al. improve on native Yao's garbled circuits by moving
more computation to the offline (input independent) phase, and thereby
achieving lower timings for the online (input dependent) phase. The
overall times remain roughly the same. As in many other works, we are
interested in the overall execution times of our
protocols -- not just the online phase. Additionally, the benchmarks
provided in Groce et al. (and also Bost et al.) are small "warm-up"
benchmarks that we provide in our paper for completeness (vector
dimensions are 30 and 47). The full power of our techniques can be
best seen in the ease of running large benchmarks such as Tensorflow
neural networks (this is one of our simplest large benchmarks, for
which no prior secure computation protocols even exist); the base
inner product for this compute is over vectors of length 784. The
Boolean circuit for this inner product itself would roughly consist of
800k gates (whereas the number of gates in the hybrid approach for the
entire computation is only 35k). The work of Groce et al. can
potentially be used as a backend  for the Boolean part of the compute
in our protocols - we leave this exploration to future work.

Secure code partitioning:

Secure code partitioning is not a new technique by itself. We only
claim that it provides a generic way to execute large programs without
going into the details of underlying individual secure protocols. We
implement it only in the arithmetic setting currently as share
reconstruction is simple addition, which has roughly the same cost in
both the Boolean and arithmetic settings.

Secure array accesses:

Combining the techniques of EzPC and ObliVM to support
Boolean, arithmetic and secret array access would be an interesting
future direction. However, for ML algorithms such as logistic
regression, linear classifier, DNNs/CNNs, the access patterns are
uniform and so providing for secret array accesses will not improve
performance.

Integrating in an existing compiler:

We found that we can write many secure prediction algorithms
in an EzPC-like simple language, without pulling in complexities of
other languages and systems. When we add more features to EzPC, such
as ORAM support, we could definitely consider the option of
integrating with some other compiler.

========== END ==========

[AR]: What is the high-level point of the rebuttal? Worth thinking about that.

[DG]: I agree with Aseem's comment above.

[NC]: I think it should be to address Rev 1's main concern that garbled circuit implementations alone might suffice as Divya says below.

[DG] Regarding the Rev #1, I feel that he is just not convinced that we need a mix of arithmetic and boolean representations for efficiency reasons. He seems to come from Yao garbled circuits club and believes that with enough optimizations Yao can work best in all contexts. And we need to contest this belief. For this, we should give citations of all papers that argued a mix of arithmetic and boolean starting with line of work of HE+Yao, ABY, SecureML, MiniONN, Gazelle and more. We should point out that even specialized protocols for NNs and other big tasks did indeed use mix of A and B. And this is critical. It has been observed by various works that doing large number of multiplications using Yao is terrible (huge size of boolean circuit resulting in big communication complexity etc).

[DG] Then regarding Groce et al. we can write that it does better than standard Yao by pushing more stuff from online to offline phase. It does no improvement in overall time of a GC based computation. And just like all other works, we are intersted in overall end-to-end time and not just online time unlike Groce et al.

We thank the reviewers for their useful comments. We address a few of them below.

First, our language does indeed allow providing input that is public,
by specifying the value to be a constant public value (is this correct
and can we write the syntax for it?). An example of public input is
loop counter value. This ensures that such values need not be
secret-shared leading to high overheads. However, this input is
hard-coded during compile time and we currently do not allow users to
input such values at runtime (can we say why?)

[DG] the reason for above is the following. Given the ABY code, the circuit generation phase happens on both client and server independently and then ABY compares these circuits for equality. If we allowed dynamic public inputs at circuit generation step, this public value would have to be communicated to the other party. Currently, it is not supported directly in ABY. There is no fundamental reason for this, just that they didnt support this.

[AR]: The reviewer is making a distinction between compile time
constants and public inputs.

[DG] I agree.

[AR] For example, a public input might be
implemented as follows:

let x = input1 () in
let y = declassify x in
//now on y is public

We can't support this currently (we had this support in Wysteria
because we were interpreting the programs).

[DG] This can be one way to implement public values by declassifying it.

We can rebut by saying: "We don't support it currently, the parties
would have to do out-of-band communication, and hardcode them in the
program. For our ML benchmarks, we haven't needed this support."

[DG] I agree.

As noted in prior works ([19], [27], [32]), purely Boolean circuits
have reasonable performance when dealing with small input sizes
(e.g. Groce et al.); even a simple task such as matrix multiplication
can be infeasible to run using garbled circuits when the matrix
dimensions are large (e.g. in our machine learning applications). In
fact, for these applications, the Yao’s circuit generated would not
fit in memory.

[DG] Above might not be valid argument with Yao pipelining technique. It was just that ABY does not have Yao pipelining supported. I have some suggestions on how to address this above.

Additionally, the work of Groce et al. only compares
online times with prior work and do not include pre-processing
time. We remark that these benchmarks are too small for any
appreciable differences in performance between various protocols and
we only included them to compare our work with other secure ML
classification works. However, we thank the reviewer for the pointer
to Groce et al. and shall include it in the paper. Another point to be
noted is that our framework can be modified to incorporate further
optimized backends (for Boolean, e.g. Groce et al.) and for
Arithmetic.

[AR]: I don't understand this well. Is this trying to say that Groce
et al.'s benchmarks are small, and so their pure Yao approach works
well, and for large benchmarks it would be terrible? The reviewer also
says that our benchmarks are simple. Can we be more forceful if indeed
Groce benchmarks are smaller than ours?

[DG] We can certainly quantify this. For example, Groce et al. runs experiments for linear classifier or hyperplane based thing where the length of inner product is 30 or 47. Compare this with simplest model for MNIST which is logistic regression, where we need to do an inner product of length 784. Also, we can look at numofgates in Tables of experiments, and it is clear that for DNNs, CNNs and matrix factorization, number of gates is just huge. Also, here we can note that for DNNs this number of gates is for mix of A and B circuit. If we write size of just boolean circuit, it would be order of magnitude bigger. Some of these arguments can be used to justify that our benchmarks are much larger and no one has done them earlier.

We do not claim that our compiler choices are optimal. Multiplication
is always done using arithmetic secret sharing as the cost of matrix
multiplication in Boolean shares is much higher; addition is done
either over Boolean or Arithmetic shares depending on prior shares
that are available and not stale. Obtaining a compiler with better
choices is an interesting optimization problem that could potentially
consider other parameters such as network latency, bandwidth, compute
resources and so on. We leave this exploration to future work.

[DG] We should quantify by saying that for 64-bit integer multiplication, using A shares, we need to do 2 multiplications of uint64. Over Yao, the circuit size would 64*64 and would require at least 64*64 AES operations in online phase.

Secure code partitioning is not a new technique by itself. We only
claim that it provides a generic way to execute large programs without
going into the details of underlying individual secure protocols. We
implement it only in the arithmetic setting currently as share
reconstruction is simple 32-bit addition, which has roughly the same
cost in both the Boolean and arithmetic settings.

Our new language is superior to other prior domain-specific language
in that it is the only language to allow compilation into a backend
that allows mixed arithmetic and Boolean computation. The task of
figuring out which part of compute to execute using which type of
sharing is unique to our compiler.

[DG] May be say something like "it is the first framework where the compiler chooses between different protocols based on cryptographic costs. All earlier DSL's in this space only had a single back-end and there was no choice to be made." 
I think this is a major contribution on the language and compiler front.

We agree with the reviewer that EzPC does not support secret array
access and combining the techniques of EzPC and ObliVM to support
Boolean, arithmetic and secret array access would be an interesting
future direction. However, for our benchmarks (e.g. ML algorithms such
as logistic regresion, linear classifier, DNNs/CNNs), the access patterns are uniform and so providing for
secret array accesses will not improve performance.
We agree with the other comments made by the reviewers and will
incorporate them accordingly. One or two comments have not been
addressed and we are slightly over the word limit (500 words).

NEW DRAFT BELOW
----------------

We thank the reviewers for their useful comments. We address a few of them.

First, our language allows providing public input, by specifying the value to be a compile-time constant value. An example is loop counter value. However, we do not currently support run-time public values to be input by parties (primarily because the underlying ABY backend does not provide support for this). Such values would have to be communicated out-of-band to the other party, and hardcoded into the program. We remark that for our ML benchmarks, we haven't needed this support but acknowledge that this would be a useful feature to integrate in the future.

A long sequence of works  have provided evidence that pure Boolean circuit computation (such as Yao's garbled circuit) does not provide good efficiency, especially when running . Indeed, even specialized protocols for executing Neural Networks and other more sophisticated ML tasks  did use a (manual) mix of Arithmetic and Boolean computation for efficiency and remark that doing large number of multiplications using Yao is terrible (due to the large size of the Boolean circuit resulting in high communication).

We thank the reviewer for pointing us to

We do not claim that our compiler choices are optimal. Multiplication is always done using arithmetic secret sharing as the cost of matrix multiplication in Boolean shares is much higher (for 64-bit integer multiplication, using arithmetic shares, we need to do 2 multiplications of uint64. Over Yao, the circuit size would be 64*64 and would require at least 4096 AES operations in the online phase.). Addition is done either over Boolean or Arithmetic shares depending on prior shares that are available and not stale. Obtaining a compiler with better choices is an interesting optimization problem that could potentially consider other parameters such as network latency, bandwidth, and compute resources. We leave this exploration to future work.

Our new language is superior to other prior domain-specific languages in that it is the only language to allow compilation into a backend that allows mixed arithmetic and Boolean computation. Ours is the first framework where the compiler chooses between different protocols based on cryptographic costs. All earlier DSL's in this space only had a single back-end and there was no choice to be made.

We agree with the other comments made by the reviewers and will incorporate them accordingly.

-----------------------------------------