### 2018-03-23 usenix18/rebuttal.txt
finalizing

We thank the reviewers for their comments. We structure the rebuttal
to address main concerns in the first ~500 words, leaving the rest to
a detailed section.

Comparison with Yao-only, Groce et al., CBMC-GC etc.:

We are not the first to observe that the Yao-only approach does not
scale well beyond examples with small number of multiplications ([18],
[26], [31], [40], [37]), especially for large-scale arithmetic
computations commonly found in Machine Learning (ML) algorithms -- the
main use case of EzPC. Indeed, this observation is one of the main
drivers behind development of the specialized MPC protocols in previous
work ([40], [37]). Therefore, it is critical to combine the
arithmetic and boolean circuits for the kind of ML algorithms we apply
EzPC to.

We thank Reviewer A for the Groce et al. reference. Firstly, our
numbers  cannot be compared directly with Groce et al. since they
report only the "online" time of the protocols, while we report the
sum of online and "offline" times. Further, the two overlapping
benchmarks in Groce et al. are quite small, and so, it is not
surprising that the Yao-only approach works well. It is not clear that
Groce et al.'s approach would scale to larger EzPC examples, e.g. DNNs
or matrix factorization (more thorough comparison below).

As for comparing with the other compilers, we go a step further and
compare with the best known previous specialized protocols,
except for benchmarks such as Bonsai, where ours is the first MPC
implementation. Compilers such as CBMC-GC have not been used at the
scale of benchmarks that we have presented in the paper.

Performance gains not being impressive:

For large benchmarks, such as CIFAR 10 (Table 7) and Matrix
Factorization (Table 8), EzPC performs much better than the previous
best known algorithms (~2x and ~19x resp.). Unsurprisingly, for small
benchmarks in Tables 1-4, where the baseline time itself is only
~5seconds, the performance gains of EzPC do not seem as
impressive. But even then we would like to point out that while most
of the previous work on those benchmarks developed specialized
protocols, EzPC framework is general and programmer friendly.

Language features:

EzPC language itself is quite simple and currently does not support
features such as secret indices and public inputs. However, the
language design has been driven by machine learning, specifically
secure prediction, case studies. As we show, even with such a simple
language, we can implement a lot of such examples.

We also remark that while the source language is simple, EzPC compiler
is the first compiler that generates a mix of arithmetic and boolean
circuits, while hiding all the low-level crypto details from the
programmer.

Cost-awareness

Our compiler currently applies certain heuristics to choose arithmetic
or boolean circuit for different parts of the program (Sec. 5). Even
with such simple heuristics, we are able to achieve the demonstrated
efficiency and scalability. A more precise cost model that factors in
the cost of network, input sizes, etc. would only improve performance
(more on it below).

==============================================

Detailed responses:

Comparison with Groce et al.:

Groce et al. improve on native Yao's garbled circuits by moving
more computation to the offline (input independent) phase, and thereby
achieving lower timings for the online (input dependent) phase. The
overall times remain roughly the same. As in many other works, we are
interested in the overall execution times of our
protocols -- not just the online phase. Additionally, the benchmarks
provided in Groce et al. (and also Bost et al.) are small "warm-up"
benchmarks that we provide in our paper for completeness (vector
dimensions are 30 and 47). The full power of our techniques can be
best seen in the ease of running large benchmarks such as Tensorflow
neural networks (this is one of our simplest large benchmarks, for
which no prior secure computation protocols even exist); the base
inner product for this compute is over vectors of length 784. The
Boolean circuit for this inner product itself would roughly consist of
800k gates (whereas the number of gates in the hybrid approach for the
entire computation is only 35k). The work of Groce et al. can
potentially be used as a backend  for the Boolean part of the compute
in our protocols - we leave this exploration to future work.

Cost aware compiler:

We do not claim that our compiler choices are optimal. Multiplication
is always done using arithmetic secret sharing as the cost of matrix
multiplication in Boolean shares is much higher (for 64-bit integer
multiplication, using arithmetic shares, we need to do 2
multiplications of uint64. Over Yao, the circuit size would be 64*64
and would require at least 4096 AES operations in the online
phase.). Addition is done either over Boolean or Arithmetic shares
depending on prior shares that are available and not stale.

Secure code partitioning:

Secure code partitioning is not a new technique by itself. We only
claim that it provides a generic way to execute large programs without
going into the details of underlying individual secure protocols. We
implement it only in the arithmetic setting currently as share
reconstruction is simple addition, which has roughly the same cost in
both the Boolean and arithmetic settings.

Secure array accesses:

Combining the techniques of EzPC and ObliVM to support
Boolean, arithmetic and secret array access would be an interesting
future direction. However, for ML algorithms such as logistic
regression, linear classifier, DNNs/CNNs, the access patterns are
uniform and so providing for secret array accesses will not improve
performance.

Integrating in an existing compiler:

We found that we can write many secure prediction algorithms
in an EzPC-like simple language, without pulling in complexities of
other languages and systems. When we add more features to EzPC, such
as ORAM support, we could definitely consider the option of
integrating with some other compiler.