### 2018-03-24 usenix18/rebuttal.txt
minor changes

We thank the reviewers for their comments. We structure the rebuttal
to address main concerns in the first ~500 words, leaving the rest to
a detailed section.

Comparison with Yao-only, Groce et al., CBMC-GC etc.:

We are not the first to observe that the Yao-only approach does not
scale well beyond examples with small number of multiplications ([18],
[26], [31], [40], [37]), especially for large-scale arithmetic
computations commonly found in Machine Learning (ML) algorithms -- the
main use case of EzPC. Indeed, this observation is one of the main
drivers behind development of the specialized MPC protocols in previous
work ([40], [37]). Therefore, it is critical to combine the
arithmetic and boolean circuits for the kind of ML algorithms we apply
EzPC to.

We thank Reviewer A for the Groce et al. reference. Firstly, our
numbers  cannot be compared directly with Groce et al. since they
report only the "online" time of the protocols, while we report the
sum of online and "offline" times. Further, the two overlapping
benchmarks in Groce et al. are quite small, and so, it is not
surprising that the Yao-only approach works well. It is not clear that
Groce et al.'s approach would scale to larger EzPC examples, e.g. DNNs
or matrix factorization (more thorough comparison below).

As for comparing with the other compilers, we go a step further and
compare with the best known previous specialized protocols,
except for benchmarks such as Bonsai, where ours is the first 2PC
implementation. Compilers such as CBMC-GC have not been used at the
scale of benchmarks that we have presented in the paper.

Performance gains not being impressive:

For large benchmarks, such as CIFAR 10 (Table 7) and Matrix
Factorization (Table 8), EzPC performs much better than the previous
best known algorithms (~2x and ~19x resp.). Unsurprisingly, for small
benchmarks in Tables 1-4, where the baseline time itself is only
~5seconds, the performance gains of EzPC do not seem as
impressive. But even then we would like to point out that while most
of the previous work on those benchmarks developed specialized
protocols, EzPC framework is general and programmer friendly.

Language features:

EzPC language itself is quite simple and currently does not support
features such as secret indices and public inputs. However, the
language design has been driven by machine learning, specifically
secure prediction, case studies. As we show, even with such a simple
language, we can implement a lot of such examples.

We also remark that while the source language is simple, EzPC compiler
is the first compiler that generates a mix of arithmetic and boolean
circuits, while hiding all the low-level crypto details from the
programmer.

Cost-awareness

Our compiler currently applies certain heuristics to choose arithmetic
or boolean circuit for different parts of the program (Sec. 5). Even
with such simple heuristics, we are able to achieve the demonstrated
efficiency and scalability. A more precise cost model that factors in
the cost of network, input sizes, etc. would only improve performance
(more on it below).

==============================================

Detailed responses:

Comparison with Groce et al.:

Groce et al. improves on online time and online communication of native Yao's 
garbled circuits by moving more computation to the offline (input independent) phase. 
However, the overall times remain roughly the same. As is standard, we are
interested in the overall execution times of our
protocols -- not just the online phase. Additionally, the benchmarks
provided in Groce et al. (and also Bost et al.) are small "warm-up"
benchmarks that we provide in our paper for completeness (vector
dimensions for inner product are 30 and 47).
The full power of our techniques can be best seen in the ease of 
running large benchmarks such as logistic regression, and neural networks. 
These require much larger number of multiplications and no one has 
attempted to implement these using garbled circuits. For instance, 
simplest benchmark of logistic regression requires 10 inner products 
of length 784. The Boolean circuit for one inner product itself would 
roughly consist of 800k gates (whereas the number of gates in the hybrid
 approach for the entire computation is only 35k). The work of Groce et al. 
can potentially be used as a backend for the Boolean part of the compute 
in our protocols - we leave this exploration to future work.

Cryptographic Cost-aware compiler:

We do not claim that our compiler choices are optimal. Multiplication
is always done using arithmetic secret sharing as the cost of
multiplication in Boolean shares is much higher (for 64-bit integer
multiplication, using arithmetic shares, we need to do only 2
multiplications of uint64. Over Yao, the circuit size would be 64*64
and would require at least 4096 AES operations in the online
phase.). Addition *does not* have a default way of evaluation and is done 
either over Boolean or Arithmetic shares depending on available shares of operands.

Secure code partitioning:

Our technique of secure code partitioning, although not entirely new, provides a 
generic way to execute large programs without going into the specific details of 
cryptographic protocol. We describe it only in the arithmetic setting but as the
reviewer correctly points out, same can be done in Boolean setting as well.

Secure array accesses:

Combining the techniques of EzPC and ObliVM to support
Boolean, arithmetic and secret array access would be an interesting
future direction. However, for ML algorithms such as logistic
regression, linear classifier, DNNs/CNNs, the access patterns are
uniform and so providing for secret array accesses will not improve
performance.

Integrating in an existing compiler:

We found that we can write many secure prediction algorithms
in an EzPC-like simple language, without pulling in complexities of
other languages and systems. When we add more features to EzPC, such
as ORAM support, we could definitely consider the option of
integrating with some other compiler.