### 2018-04-27 ccs/conclude.tex
starting ccs repo

%% \section{Conclusion}
%% \label{sec:conclude}
%% We presented \tool, the first
%% cryptographic-cost aware framework that generates efficient and
%% scalable \mpc protocols from high-level programs.
%% The generated protocols comprise combinations
%% of arithmetic and boolean circuits and have performance comparable to, or better than the custom, specialized
%% protocols from previous works.
%% The \mpc backend we use provides security against semi-honest or passive adversaries.
%% %our compiler is oblivious to the ``type of adversary'' for the underlying \mpc backend.
%% Given a \mpc backend secure against malicious or active adversaries, Theorem \ref{theorem:security} 
%% can be extended to obtain a framework with malicious security. Furthermore,  
%% partitioning can be modified to provide malicious security by {\it signing} the shares of the intermediate states. However, we are not aware of efficient maliciously secure \mpc implementation for combinations of
%% arithmetic and boolean circuits and hence leave a complete treatment of malicious security to future work.
%% % Several areas of improvement remain: a) While the backend that we use~\cite{aby} only provides security against a semi-honest adversary, our compiler is oblivious to the ``type of adversary'' for the underlying backend cryptographic protocol. In particular, if our cryptographic backend were maliciously secure, then Theorem \ref{theorem:security} would provide us a framework with malicious security. Furthermore, modifying our code partitioning to provide malicious security is fairly easy (one would instead provide {\em signed} shares of intermediate states in the decomposed programs to obtain malicious security). b) Our compiler converts programs into circuits before executing a \mpc and is efficient for the applications we present. However, one could consider executing programs that exploit secret-dependent memory accesses (such as binary search) to obtain better performance (as in \cite{oblivm}). It would be interesting to see if our techniques can be combined with that of \cite{oblivm}. \nc{Too much? Should we simply say only about the malicious security and stop?}