### 2018-07-29 ndss/main.tex
added ndss version.

\documentclass[conference]{IEEEtran}

\pagestyle{plain}
\input{headers}

\usepackage{epsfig,endnotes,pdfpages}
\usepackage{wrapfig}
\usepackage{float}

\begin{document}

\title{\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation for Machine Learning}

%\setlength\intextsep{1pt}
%\setlength{\textfloatsep}{1pt}
%\setlength{\abovecaptionskip}{4pt}
%\setlength{\belowcaptionskip}{4pt}

\maketitle

\begin{abstract}
We present \tool: a 
%new ``cryptographic-cost aware'' 
secure two-party computation (\mpc) framework that generates efficient
\mpc protocols from high-level, easy-to-write programs.
\tool provides formal correctness and security guarantees while maintaining performance and scalability.
%% that 
%% %contain absolutely no cryptographic details. Programs in our language 
%% are easy-to-write and are statically verified. 
%Furthermore,
Previous language frameworks, such as CBMC-GC, ObliVM, SMCL, and
Wysteria,
generate protocols that use either arithmetic or boolean
circuits exclusively. Our compiler is the first to generate protocols
that combine both arithmetic sharing and garbled circuits  for better
performance.
%We give the first compiler that generates protocols which combine both arithmetic and boolean circuits
%for better performance compared to protocols that use either arithmetic or boolean circuits exclusively.
We
empirically demonstrate that the
protocols generated by our framework match or outperform (up to
19x)
recent works that provide hand-crafted protocols for various
functionalities such as secure  prediction  and matrix factorization.

\end{abstract}

\graphicspath{{./Images/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\input{intro}
%\input{fig-ex-sml}
%\input{fig-ex-aby}

\input{ex}
\input{ld}
\input{security}
\input{pipe}
\input{impl}

\input{eval}
\input{train}
\input{related}
\input{conclude}

%\newpage

\bibliographystyle{IEEEtranS}
\bibliography{bib}

%\bibliographystyle{abbrv}
%\bibliography{bib}
%\balance

\appendix

\input{appendix_b}

\input{appendix_c}

\input{appendix_a}

\end{document}